Space-time plots

### 2016-06-02 Rmd/space-time.Rmd
Full draft as of 1 June 2016

# Space versus time analysis
```{r, message = FALSE, warning=FALSE, results='hold'}
library(ecoscales)
library(readxl)
library(stringdist)
library(maptools)

p_root <- set_base_path("ecoscales")
p_dat <- full_path(p_root, "external/data/result")
p_calib <- full_path(p_root, "external/data/calibration")
# dat <- as.data.table(read_excel(full_path(p_dat, "merged.xlsx")))
dat <- as.data.table(read_excel(full_path(p_dat, "mergedf.xlsx")))
cal <- data.table(read_excel(fp(p_calib, "merged_calibrationff.xlsx")))

setnames(dat, "DOI/title", "DOI")
dat[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
dat[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace

dat <- dat[, names(dat)[1:20], with = FALSE]
```

`r length(unique(dat$DOI))` papers  main analysis

## Data prep
### Merge full and calibration datasets
```{r, message=FALSE, warning=FALSE, results='hide'}
dat2 <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
                     t_btwn_samp, study_duration)])

rnms <- c("DOI", "feature", "observer", "plot_res", "n_sites", 
          "sampled_area", "t_btwn_samp", "study_duration")
calr <- cal[, rnms, with = FALSE]
for(j in names(calr)[c(2, 4:8)]) set(calr, j=j, value = as.numeric(calr[[j]]))

# select down to columns of interest
calr <- calr[feature != 99, lapply(.SD, mean), by = .(DOI, feature), 
             .SDcols = rnms[4:8]]
length(unique(calr$DOI)) + length(unique(dat$DOI))  # 148 papers providing dat

calr[, samp_duration := NA]
calr[, c("DOI", "feature") := NULL]
setcolorder(calr, names(calr)[c(1:3, 6, 4:5)])

datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))
```

@@

## Analyses

### Resampling with uncertainties

@@

```{r, eval=FALSE}
p <- c(0.64, 0.84, 0.53, 1.21)  # from calibration.Rmd

@@

### Bootstrapping

Good old-fashioned bootstrapping with replacement, no perturbation of value
```{r, eval = FALSE}
# bootstrap
set.seed(1)
bootsamp <- lapply(1:1000, function(x) {
  if((x / 100) %in% 1:100) print(x)
  dnew <- copy(datf[, .(plot_res, sampled_area, t_btwn_samp, study_duration)])
  dnew <- dnew[sample(1:.N, size = .N, replace = TRUE), ]
  dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
  dnew
})

@@

@@

# labels for temporal axes
# return interval
tlab1 <- c(expression(paste(NULL<="second"^-1)), expression("minute"^-1),
           expression("hour"^-1), expression("day"^-1), 
           expression("week"^-1), expression("month"^-1),            
           expression("year"^-1), expression("decade"^-1),
           expression("century"^-1), 
           expression("millenium"^-1), "once-off")# expression(infinity))
taxis1 <- cbind.data.frame( 
  c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 7, 30, 365, 365 * 10, 
    365 * 100, 365 * 1000, 365 * 10000))
taxis1 <- data.table(taxis1)
setnames(taxis1, names(taxis1), "days")
taxis1[, logdays := log10(days)]

@@

cxa = 0.7
pdf("paper/figures/res_v_extent_pbstrap.pdf", width = 7, 
    height = 2.5)
par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
    mgp = c(2, 0.5, 0))
# image(rres, col = cols, breaks = brks,
image(rres, col = rev(terrain.colors(30)), breaks = brks,
     axes = FALSE, xlab = "", ylab = "")
# plot(rres2, col = rev(terrain.colors(20)))
axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
# plot(rres2, axes = FALSE, legend.only = TRUE)
aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
image(rext, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
      ylab = "", breaks = brks2)
axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
     tcl = -0.2)
aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
              cex.axis = cxa, tcl = -0.1)
plot(rext, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
dev.off()

### Bootstrap
```{r, eval = FALSE}
bootdt <- rbindlist(bootsamp)
resdat2 <- data.table(x = log10(bootdt$t_btwn_samp), y = log10(bootdt$plot_res))
resdat2$n <- rep(1, nrow(resdat2))
rx <- range(resdat2$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
rx[2] <- 6.56
resdat2[x < rx[1], x := rx[1]]
resdat2[x > rx[2], x := rx[2]]
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat2[y < ry[1], y := ry[1]]
resdat2[y > ry[2], y := ry[2]]
coordinates(resdat2) <- ~x + y

@@

cxa = 0.7
pdf("paper/figures/res_v_extent_bstrap.pdf", width = 7, 
    height = 2.5)
par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
    mgp = c(2, 0.5, 0))
# image(rres, col = cols, breaks = brks,
image(rres2, col = rev(terrain.colors(30)), breaks = brks,
     axes = FALSE, xlab = "", ylab = "")
# plot(rres2, col = rev(terrain.colors(20)))
axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
# plot(rres2, axes = FALSE, legend.only = TRUE)
aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))

@@

@@

# densities
# resolution
rres3 <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat3, 1)
rres3 <- (rres3 / cellStats(rres3, sum)) * 100
rng <- range(rres3[is.finite(rres3)])
bwidth <- (rng[2] - rng[1]) / 30
brks <- seq(rng[1], rng[2], bwidth)
brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)

# extent
rext3 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat3, 1)
rext3 <- (rext3 / cellStats(rext3, sum)) * 100
rng <- range(rext3[is.finite(rext3)])
# rng[2] <- 0.1
bwidth <- (rng[2] - rng[1]) / 30
brks2 <- seq(rng[1], rng[2], bwidth)
brklabs2 <- brklabs #round(brks2[seq(1, length(brks2), 3)], 2)

@@

# cxa = 0.7
ll <- 3.5
cxa = 0.8
# pdf("paper/figures/res_v_extent.pdf", width = 7, height = 2.5)
pdf("paper/figures/kde4.pdf", width = 7, height = 5.5)
par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
    mgp = c(2, 0.25, 0))
# resolutions
# image(rres3, col = rev(terrain.colors(30)), breaks = brks,
image(rres3, col = rev(terrain.colors(50)), 
     axes = FALSE, xlab = "", ylab = "")
axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Sampling interval", 1, cex = cxa, line = ll)
mtext("Spatial resolution", 2, cex = cxa, line = ll)
aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     #legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
     legend.shrink = 0.9, col = rev(terrain.colors(50)))

# extents
# image(rext3, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
image(rext3, col = rev(terrain.colors(50)), axes = FALSE, xlab = "", 
      ylab = "")#, breaks = brks2)
axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
     tcl = -0.2)
aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
              cex.axis = cxa, tcl = -0.1)
mtext("Temporal duration", 1, cex = cxa, line = ll)
mtext("Spatial extent", 2, cex = cxa, line = ll)
plot(rext3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     #legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
     legend.shrink = 0.9, col = rev(terrain.colors(50)))

# spatial
image(sres, col = rev(terrain.colors(100)), axes = FALSE, xlab = "", ylab = "")
# points(spatdat$x, spatdat$y, cex = 0.1)
axis(1, at = aaxis1$logres, labels = alab1, las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Spatial resolution", 1, cex = cxa, line = ll)
mtext("Spatial extent", 2, cex = cxa, line = ll)
aargs <- list(mgp = c(3, 0.25, 0), at = seq(0, 0.12, 0.02), 
              labels = seq(0, 0.12, 0.02), cex.axis = cxa, tcl = -0.1)
plot(sres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = rev(terrain.colors(100)))

# temporal
image(tres, col = rev(terrain.colors(50)), axes = FALSE, xlab = "", ylab = "")
# points(spatdat$x, spatdat$y, cex = 0.1)
axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
mtext("Sampling interval", 1, cex = cxa, line = ll)
mtext("Temporal duration", 2, cex = cxa, line = ll)
aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
              cex.axis = cxa, tcl = -0.1)
plot(tres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = rev(terrain.colors(100)))
dev.off()

@@

## Ancillary/supplementary plots and stats

### Statistics
```{r, eval=FALSE}
length(which(dat$study_type %in% 
               c("remote sensing", "other geographic data"))) / nrow(dat)
dat[t_btwn_samp == 0, .N] / dat[, .N]  # 31% of samples are once-offs
dat[study_duration == samp_duration, .N] / nrow(dat)

@@

pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
par(mfrow = c(1, 4), mar = c(4, 2, 1, 1), oma = c(2, 3, 0, 0))
# resolution
h <- hist(hdat$res, breaks = aaxis1$logres, plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", 
     xlab = "Spatial resolution", ylim = c(0, 25), main = "", col = "red1")
axis(1, at = aaxis1$logres, labels = alab1, las = 2, cex.axis = cxa, 
     tcl = -0.2, mgp = c(2, 0.25, -0.45))
mtext("Percent of observations", side = 2, line = 3, cex = 0.8)

# extent
h <- hist(hdat$ext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", 
     xlab = "Spatial extent", ylim = c(0, 20), main = "", col = "red3")
axis(1, at = c(-6, aaxis2$logarea), 
     labels = c(expression(paste(NULL<="0.01m"^2)), expression(paste(0.1, m^2)), 
                alab2[-1]), las = 2, cex.axis = cxa, 
     tcl = -0.2, mgp = c(2, 0.25, -0.45))

# interval
tax <- c(taxis2$logdays[-c(10:11)], mean(taxis2$logdays[c(10:11)]))
h <- hist(hdat$int, breaks = taxis1$logdays, plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", 
     xlab = "Sampling interval", ylim = c(0, 30), main = "", col = "blue1")
axis(1, at = tax, labels = tlab1[-10], las = 2, cex.axis = cxa, 
     tcl = -0.2, mgp = c(2, 0.25, -0.45))

>>>>

# Space versus time analysis
```{r, message = FALSE, warning=FALSE, results='hold'}
library(ecoscales)
library(readxl)
library(stringdist)
library(maptools)
library(RColorBrewer)

p_root <- set_base_path("ecoscales")
p_dat <- full_path(p_root, "external/data/result")
p_calib <- full_path(p_root, "external/data/calibration")
# dat <- as.data.table(read_excel(full_path(p_dat, "merged.xlsx")))
dat <- as.data.table(read_excel(full_path(p_dat, "mergedfff.xlsx")))
cal <- data.table(read_excel(fp(p_calib, "merged_calibrationffff.xlsx")))
full <- data.table(read_excel(fp(p_dat, "full_less_Treuer.xlsx")))

setnames(dat, "DOI/title", "DOI")
dat[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
dat[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
full <- gsub("DOI:|doi:|DOI: |doi: |DOI ", "", full$DOI)
full <- gsub("\\s", "", full)  # remove DOI and whitespace

dat <- dat[, names(dat)[1:20], with = FALSE]

unique(dat$study_type)
dat[, study_type := tolower(study_type)]
kwords <- c("field", "paleo", "remote", "automated", "other")
for(i in 1:length(kwords)) dat[like(study_type, kwords[i]), st := i]

# dat[st == 5, .N] / nrow(dat)

# class(cal$DOI)
# class(dat$DOI)
# unique(cal$DOI) %in% unique(dat$DOI)
# match(cal$DOI, dat$DOI)
# match(dat$DOI, cal$DOI)
# dat[dat$DOI %in% cal$DOI, DOI]
# dat[cal$DOI %in% dat$DOI, ]
# cal[DOI == dat[cal$DOI %in% dat$DOI, DOI][11]]
# dat[DOI == dat[cal$DOI %in% dat$DOI, DOI][7]]

## Data prep
### Merge full and calibration datasets
```{r, message=FALSE, warning=FALSE, results='hide'}
dat2 <- copy(dat[, .(st, DOI, study_type, plot_res, n_sites, sampled_area,
                     samp_duration, t_btwn_samp, study_duration)])
rnms <- c("st", "DOI", "study_type", "feature", "observer", "plot_res", 
          "n_sites", "sampled_area", "t_btwn_samp", "study_duration")
calr <- cal[, rnms, with = FALSE]
for(j in names(calr)[c(4, 6:10)]) set(calr, j=j, value = as.numeric(calr[[j]]))

# select down to columns of interest
# st <- calr$study_type[which(calr$feature != 99)]
calr <- calr[feature != 99, lapply(.SD, mean), by = .(DOI, feature), 
             .SDcols = rnms[c(1, 6:10)]]
length(unique(calr$DOI)) + length(unique(dat$DOI))  # 140 papers providing dat
unique(calr$DOI)
calr_type <- data.table(cbind("DOI" = unique(calr$DOI), # calib field studies
                              "study_type" = c(rep("field", 5), "paleo",
                                               rep("field", 4))))

# combine full and calibration
calr[, samp_duration := NA]
calr[, "feature" := NULL] #c("DOI", "feature") := NULL]
calr <- merge(calr, calr_type, by = "DOI")
setcolorder(calr, names(calr)[c(2:1, 9, 3:5, 8, 6:7)])
datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))

doisup <- c(unique(cal$DOI_data_source), unique(dat$DOI_data_source))
doisup <- doisup[!is.na(doisup) & doisup != "NA" & doisup != "-"]
doisup <- unlist(strsplit(doisup, ";"))

`r length(unique(datf$DOI))` papers  main analysis

`r length(unique(full)) + 20 + 57` papers main analysis (everyone's but Tim's)

`r (length(unique(full)) + 20 + 57) / 42918 * 100` % of all papers since 2004

`r length(unique(doisup))` additional papers or other publications tracked down

`r nrow(datf)` records (ecological observations)

<a href="#top">Back to top</a>

@@

<a href="#top">Back to top</a>

## Analyses
### Resampling with uncertainties

@@

```{r, eval=FALSE}
# p <- c(0.64, 0.84, 0.53, 1.21)  # from calibration.Rmd
p <- c(0.58, 0.82, 0.99, 1.24)  # from calibration.Rmd

@@

<a href="#top">Back to top</a>

```{r, eval = FALSE, echo = FALSE}
### Bootstrapping
# Good old-fashioned bootstrapping with replacement, no perturbation of value
# bootstrap
set.seed(1)
bootsamp <- lapply(1:1000, function(x) {
  if((x / 100) %in% 1:100) print(x)
  dnew <- copy(datf[, .(plot_res, sampled_area, t_btwn_samp, study_duration)])
  dnew <- dnew[sample(1:.N, size = .N, replace = TRUE), ]
  dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
  dnew
})

@@

<a href="#top">Back to top</a>

@@

# labels for temporal axes
# return interval
tlab1 <- c(expression(paste(NULL<="second"^-1)), expression("minute"^-1),
           expression("hour"^-1), expression("day"^-1), 
           expression("week"^-1), expression("month"^-1),            
           expression("year"^-1), expression("decade"^-1),
           expression("century"^-1), 
           expression("millenium"^-1), "unreplicated")# expression(infinity))
taxis1 <- cbind.data.frame( 
  c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 7, 30, 365, 365 * 10, 
    365 * 100, 365 * 1000, 365 * 10000))
taxis1 <- data.table(taxis1)
setnames(taxis1, names(taxis1), "days")
taxis1[, logdays := log10(days)]

@@

brkfun <- function(ext, ival, n) {
  rng <- range(ext[is.finite(ext)])
  bwidth <- (rng[2] - rng[1]) / n
  brks <- seq(rng[1], rng[2], bwidth)
  brklabs <- seq(0, round(rng[2], 2), ival)
  list("brks" = brks, "labs" = brklabs)
}
cuts <- 80

cxa = 0.7
pdf("paper/figures/res_v_extent_pbstrap2.pdf", width = 7, 
    height = 2.5)
par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
    mgp = c(2, 0.5, 0))
# image(rres, col = cols, breaks = brks,
brks <- brkfun(rres, 0.01, cuts)  # breaks
image(rres, col = rev(terrain.colors(cuts)), #breaks = brks,
     axes = FALSE, xlab = "", ylab = "")
# plot(rres2, col = rev(terrain.colors(20)))
axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
# plot(rres2, axes = FALSE, legend.only = TRUE)
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
brks <- brkfun(rext, 0.01, cuts)  # breaks
image(rext, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
      ylab = "")
axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
     tcl = -0.2)
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(rext, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
dev.off()
```

<a href="#top">Back to top</a>

```{r, eval = FALSE, echo = FALSE}
### Bootstrap - deprecated for now

bootdt <- rbindlist(bootsamp)
resdat2 <- data.table(x = log10(bootdt$t_btwn_samp), y = log10(bootdt$plot_res))
resdat2$n <- rep(1, nrow(resdat2))
rx <- range(resdat2$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
rx[2] <- 6.56
resdat2[x < rx[1], x := rx[1]]
resdat2[x > rx[2], x := rx[2]]
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat2[y < ry[1], y := ry[1]]
resdat2[y > ry[2], y := ry[2]]
coordinates(resdat2) <- ~x + y

@@

cxa = 0.7
pdf("paper/figures/res_v_extent_bstrap2.pdf", width = 7, 
    height = 2.5)
par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
    mgp = c(2, 0.5, 0))
# image(rres, col = cols, breaks = brks,
image(rres2, col = rev(terrain.colors(30)), breaks = brks,
     axes = FALSE, xlab = "", ylab = "")
# plot(rres2, col = rev(terrain.colors(20)))
axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
# plot(rres2, axes = FALSE, legend.only = TRUE)
aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))

@@

# datf[, plot(log10(study_duration), log10(sampled_area), xlim)]
```

<a href="#top">Back to top</a>

@@

# densities
# resolution
rres3 <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat3, 1)
rres3 <- (rres3 / cellStats(rres3, sum)) * 100
# rng <- range(rres3[is.finite(rres3)])
# bwidth <- (rng[2] - rng[1]) / 30
# brks <- seq(rng[1], rng[2], bwidth)
# brklabs <- seq(0, 0.12, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
# brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)

# extent
rext3 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat3, 1)
rext3 <- (rext3 / cellStats(rext3, sum)) * 100
# rng <- range(rext3[is.finite(rext3)])
# # rng[2] <- 0.1
# bwidth <- (rng[2] - rng[1]) / 30
# brks2 <- seq(rng[1], rng[2], bwidth)
# brklabs2 <- brklabs #round(brks2[seq(1, length(brks2), 3)], 2)

@@

brkfun <- function(ext, ival, n) {
  rng <- range(ext[is.finite(ext)])
  bwidth <- (rng[2] - rng[1]) / n
  brks <- seq(rng[1], rng[2], bwidth)
  brklabs <- seq(0, round(rng[2], 2), ival)
  list("brks" = brks, "labs" = brklabs)
}

# cxa = 0.7
ll <- 3.5
cxa = 0.8
stps <- sapply(c(3, 4), function(x) which(datf$st == x))  # IDs rs/auto/paleo
pchs <- list(1, "+")
cexs <- c(0.4, 0.6)
cuts <- 80

# pdf("paper/figures/res_v_extent.pdf", width = 7, height = 2.5)
pdf("paper/figures/kde43.pdf", width = 7, height = 5.5)
par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
    mgp = c(2, 0.25, 0))
# resolutions
# image(rres3, col = rev(terrain.colors(30)), breaks = brks,
image(rres3, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
      ylab = "")
axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Sampling interval", 1, cex = cxa, line = ll)
mtext("Spatial resolution", 2, cex = cxa, line = ll)
brks <- brkfun(rres3, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     #legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05)
# for(i in 1:length(stps)) {
#   points(resdat3$x[stps[[i]]], resdat3$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }

# extents
# image(rext3, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
image(rext3, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
      ylab = "")#, breaks = brks2)
axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
     tcl = -0.2)
brks <- brkfun(rext3, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
mtext("Temporal duration", 1, cex = cxa, line = ll)
mtext("Spatial extent", 2, cex = cxa, line = ll)
plot(rext3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     #legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05)
# for(i in 1:length(stps)) {
#   points(extdat3$x[stps[[i]]], extdat3$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }

# spatial
image(sres, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
      ylab = "")
# points(spatdat$x, spatdat$y, cex = 0.1)
axis(1, at = aaxis1$logres, labels = alab1, las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Spatial resolution", 1, cex = cxa, line = ll)
mtext("Spatial extent", 2, cex = cxa, line = ll)
brks <- brkfun(sres, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(sres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
mtext("C", side = 3, line = -1, cex = 0.8, adj = 0.05)
# for(i in 1:length(stps)) {
#   points(spatdat$x[stps[[i]]], spatdat$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }

# temporal
image(tres, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", ylab = "")
# points(spatdat$x, spatdat$y, cex = 0.1)
axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
mtext("Sampling interval", 1, cex = cxa, line = ll)
mtext("Temporal duration", 2, cex = cxa, line = ll)
brks <- brkfun(tres, 0.02, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(tres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
mtext("D", side = 3, line = -1, cex = 0.8, adj = 0.05)
# for(i in 1:length(stps)) {
#   points(tempdat$x[stps[[i]]], tempdat$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }
dev.off()

@@

# image(rres3, col = rev(terrain.colors(100)))
# points(resdat3$x, resdat3$y, pch = 20, cex = 0.5)
image(sres, col = rev(terrain.colors(100)))
points(spatdat$x, spatdat$y, pch = 20, cex = 0.5)
```

<a href="#top">Back to top</a>

## Stats and supplementary plots

### Statistics
```{r, eval=FALSE}
length(which(dat$study_type %in% 
               c("remote sensing", "other geographic data"))) / nrow(dat)
length(which(dat$study_type %in% 
               c("remote sensing", "other geographic data"))) / nrow(dat)
dat[t_btwn_samp == 0, .N] / dat[, .N]  # 35% of samples are once-offs
dat[study_duration == samp_duration, .N] / nrow(dat)
dat[(study_duration == samp_duration) & t_btwn_samp != 0]
dat[(study_duration != samp_duration) & t_btwn_samp == 0]

@@

mga <- -0.34
sfigl <- -1.2
cxa <- 0.8
cxl <- 0.9
reds <- brewer.pal(9, name = "Reds")
blues <- brewer.pal(9, name = "Blues")
plot(1:2, pch = 20, col = blues[c(4, 6)])
yl <- 35
# pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
pdf("paper/figures/hists2.pdf", width = 5, height = 5)
par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
# resolution
h <- hist(hdat$res, breaks = aaxis1$logres, plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
     tcl = -0.2, cex.axis = cxa, xlab = "Spatial resolution", 
     ylim = c(0, yl - 10), main = "", col = reds[4], cex.lab = cxl)
axis(1, at = aaxis1$logres, labels = alab1, las = 2, cex.axis = cxa, 
     tcl = -0.2, mgp = c(2, 0.25, mga))
mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
mtext("A", side = 3, line = sfigl, cex = 0.8, adj = 0.05)

# extent
h <- hist(hdat$ext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
     tcl = -0.2, cex.axis = cxa, xlab = "Spatial extent", ylim = c(0, yl - 10), 
     main = "", col = reds[6], cex.lab = cxl)
axis(1, at = c(-6, aaxis2$logarea), 
     labels = c(expression(paste(NULL<="0.01m"^2)), expression(paste(0.1, m^2)), 
                alab2[-1]), las = 2, cex.axis = cxa,
     tcl = -0.2, mgp = c(2, 0.25, mga))
mtext("B", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
